---
title: "Technical Test"
author: "Alfonso Noguer"
date: "1/10/2019"
output:
 html_document:
   keep_md: TRUE
   theme: cerulean
   toc: true
   toc_depth: 3
   toc_float:
     collapsed: true
     smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Distance

Create an SQL query that shows the distance between the courier starting position and thepickup point, as well as the distance between the pickup point and the delivery point.

We have two tables One with the origin and another with the destinations.

For ansering all the distances we would need 3 positions and we only have two so we can't calculate both distances.

First I will program the proces in R.


### R

To calculate the distance I have two options one is directly join all 2 dataframes and combine the columns to get a distance, but that would take an awfull amount of time and memory so I discard that option.

The secon proach is to filter the 4 coordinates long/lat of origin and destinations and calculate the distance for those.

To do this we use the ID as the Key for both tables so we can know how to join origin and destination.

```{r eval=FALSE}

pacman::p_load(tidyverse)

#if we want to filter by something diferent than ID we use somethin like

key <- order_points %>% filter(order_id == "id we want to search") %>% select(id)

x1 <- orders %>% filter(id == "id/key we want to search") %>% select(acceptance_latitude)

y1 <- orders %>% filter(id == "id/key we want to search") %>% select(acceptance_longitude)

x2 <- order_points %>% filter(id == "id/key we want to search") %>% select(latitude)

y2 <- order_points %>% filter(id == "id/key we want to search") %>% select(longitude)

length_manhattan <- abs(x1-x2) + abs(y1-y2) 
# or
length_euclidean <- (x1-x2)^2 + (y1-y2)^2
```

### SQL

In SQL we do something similar first we create 2 new tables with the filtered values we care about.

Then we join both of them and calculate the distance in the select

```{SQL eval=FALSE}
SELECT a.id, a.courier_id , a.point_type, 
  SQR(POWER(a.acceptance_latitude - b_latitude,2) + 
        POWER(a.acceptance_longitude - b.longitude,2) AS distance
WHERE "contidion"
FROM orders AS a, order_points AS b
```

Being the condition whatever you want like **a.courier_id = "number"** or 
**a.coustomer_id = "number"**

## 1. Cohort Retention

Build one SQL query to create a cohort of Signup to First Order and show the result.
The objective of this cohort is to see, out of the users that signed up in Week N, how many did their first order in Week N+1, N+2, N+3...

### R

To do this in R we generate a function that retunrs the unique users in a time period in any of our dataframes.

With that function I do a double loop the first loop gives us the respective cohort, and the second calculates the retention rate by each succesive time period.
to make ir easyer to process al this process is in a function where the imputs are the two dataframes of the problem, the starting date in which we want the process to start (by default the first of january 2015 the year Glovo lauched)
and the interval of the cohorts in days (default 7 days, 1 week as the question indicates).

```{r eval=FALSE}

pacman::p_load(lubridate,tidyverse)
#frist we generate a function that finds the date_time in the dataframe 
#and returns the unique coustomer_ids if existing and costomer ids if not
#I could add a condition that detects with column is the time_stamp and selects
#
users_period_of_time <- function(orders, start_date= dmy(01-01-2015),
                                time_increment_in_days = 7){
  return(
    orders %>% 
      filter(timestamp > start_date &
               timestamp < (start_date + time_increment_in_days)) %>% 
      select(ifelse(customer_id %in% names(orders),"customer_id","id")) %>% 
      unique())
}

retention <- function(users, orders, start_date= dmy(01-01-2015),
                      time_increment_days = 7){
  last_date <- max(orders$timestamp)
  date <- start_date
  results <- list()
  while (date + time_increment_in_days < last_date) {
    users <- users_period_of_time(orders = users,
                                  start_date = date,
                                  time_increment_in_days = time_increment_days)
    total_users <- length(users)
    retention <- 0
    period <- 0
    vector <- c()
    while (retention < 1 |
           date + time_increment_in_days * period > last_date)
      {
      uniq_users_period <- users[users %in% users_period_of_time(orders = orders,
                                                        start_date = time_segment)]
      interesting_users <- users[users %in% uniq_users_period]
      users <- users[!users %in% uniq_users_period]
      retention <- retention + length(interesting_users)/total_users
      vector <- c(vector,paste0("rentention ",round(retention,2)," in period ",period))
      period = period + 1
    }#close second while
    row_name <- paste0("cohort_",date)
    results[[row_name]] <- c(paste0("total users ",total_users,vector))
    date <- date + time_increment_in_days
  }#close first while
  return(results)
}
```

### SQL

First we bucket them into different cohort by their sign up month, and store into **cohort_items**.
After that, we build **user_activities** which would return all the pairs of **(user_id, week_number)** that indicates if a user is active in that month after their original signup date.
The **WEEK_DIFF** is a function that takes in 2 dates, and return the number of months between them.
**Cohort Size:** is simply how many users are in each group.
And finally I put everithing together.

```{SQL eval=FALSE}
WITH cohort_items AS (
  SELECT
    date_trunc('week', O.activation_time_local)::date as cohort_week,
    cutomer_id
  FROM orders O
  order by 1, 2
)

WITH user_activities AS (
  SELECT
    A.user_id,
    WEEK_DIFF(
      date_trunc('week', U.timestamp)::date,
      C.cohort_week
    ) AS week_number
  FROM users U
  LEFT JOIN cohort_items C ON U.id = C.customer_id
  GROUP BY 1, 2
)

WITH cohort_size AS (
  select cohort_week, count(1) AS num_users
  FROM cohort_items
  GRUP BY 1
  ORDER BY 1
)

WITH retention_table AS (
  SELECT
    C.cohort_week,
    A.week_number,
    count(1) AS num_users
  FROM user_activities A
  LEFT JOIN cohort_items C ON A.user_id = C.user_id
  GROUP BY 1, 2
)

SELECT
  B.cohort_week,
  S.num_users AS total_users,
  B.weeek_number,
  B.num_users::float * 100 / S.num_users AS percentage
FROM retention_table B
LEFT JOIN cohort_size S ON B.cohort_week = S.cohort_week
WHERE B.cohort_week IS NOT NULL
ORDER BY 1, 3

```
